`{peacesciencer}`: An R Package for Quantitative Peace Science Research

### 2022-01-29 manuscript/ms.Rmd
update, mostly Fira Code
gonna eventually send this to Sage.

This data feature tackles a recurring problem for researchers in the peace science community. True research reproducibility is best achieved creating data from scratch, though no published guide exists that informs researchers how to do this on their own. Instead, researchers may end up reusing old code generated in past studies, leaving them to spend time and energy adjusting the sample of states, the temporal domain, and doing whatever additional troubleshooting may arise from this practice. Researchers may additionally spend too much time reproducing old code for standard information that goes into any dyadic or monadic analysis---like contiguity relationships and democracy---and have to do additional troubleshooting for how these various data sources treat missing data or treat state codes in a manner inconsistent with the more accessible Correlates of War or Gleditsch-Ward state codes. This is all compounded by changes in technology that treat the creation of the data and the analysis of data as a continuous process in which the contemporary quantitative political scientist is increasingly becoming a computer programmer as well. Graduate students and other beginners in the field face unique challenges associated with these developments. Students just learning peace science must learn how scholarship informs data in peace science and how data inform scholarship at the same time they are needing to learn quantitative methods in a chosen software package.

@@

This function has three arguments denoting the leader system to inform the creation of the leader-year data (i.e. Archigos) and two other arguments: `standardize` and `subset_years`. The `standardize` arguments, which when it defaults to "none", returns all leader-years as presented in the raw Archigos data. Archigos' leader data are nominally denominated in the Gleditsch-Ward state system data, if not necessarily Gleditsch-Ward state system dates (e.g. Archigos often has leader entries prior to state system entry in a few cases). Thus, the user can standardize leader-year data to Gleditsch-Ward state system dates of CoW state system dates. Finally, the user can subset the leader-year data to a more narrow temporal domain that may interest them with the `subset_years` function. If a user is interested in creating a data set of leader-years for an analysis of inter-state conflict initiation with the GML MID data, they can create the data that interests them with the following function. Notice how `create_leaderyears()` also returns information about the leader too, like the leader's approximate age that year, their gender, and information about their tenure.

@@

After creating the base data of interest to their project, researchers can begin to add information they want with the suite of functions outlined in Table \@ref(tab:tab-addfunctions). For example, researcher interested in a dyad-year analysis of inter-state disputes can create a non-directed dyad-year data set from 1816 to 2010 in `{peacesciencer}` with the `create_dyadyears()` function, one of the aforementioned "create" function. The following would create the base data of interest to the user (i.e. all non-directed dyad-years from 1816 to 2010).

>>>>

This manuscript tackles a recurring problem for researchers in the peace science community. True research reproducibility is best achieved creating data from scratch, though no published guide exists that informs researchers how to do this on their own. Instead, researchers may end up reusing old code generated in past studies, leaving them to spend time and energy adjusting the sample of states, the temporal domain, and doing whatever additional troubleshooting may arise from this practice. Researchers may additionally spend too much time reproducing old code for standard information that goes into any dyadic or monadic analysis---like contiguity relationships and democracy---and have to do additional troubleshooting for how these various data sources treat missing data or treat state codes in a manner inconsistent with the more accessible Correlates of War or Gleditsch-Ward state codes. This is all compounded by changes in technology that treat the creation of the data and the analysis of data as a continuous process in which the contemporary quantitative political scientist is increasingly becoming a computer programmer as well. Graduate students and other beginners in the field face unique challenges associated with these developments. Students just learning peace science must learn how scholarship informs data in peace science and how data inform scholarship at the same time they are needing to learn quantitative methods in a chosen software package.

@@

This function has three arguments denoting the leader system to inform the creation of the leader-year data (i.e. Archigos) and two other arguments: `standardize` and `subset_years`. The `standardize` argument, which when it defaults to "none", returns all leader-years as presented in the raw Archigos data. Archigos' leader data are nominally denominated in the Gleditsch-Ward state system data, if not necessarily Gleditsch-Ward state system dates (e.g. Archigos often has leader entries prior to state system entry in a few cases). Thus, the user can standardize leader-year data to Gleditsch-Ward state system dates of CoW state system dates. Finally, the user can subset the leader-year data to a more narrow temporal domain that may interest them with the `subset_years` function. If a user is interested in creating a data set of leader-years for an analysis of inter-state conflict initiation with the GML MID data, they can create the data that interests them with the following function. Notice how `create_leaderyears()` also returns information about the leader too, like the leader's approximate age that year, their gender, and information about their tenure.

@@

After creating the base data of interest to their project, researchers can begin to add information they want with the suite of functions outlined in Table \@ref(tab:tab-addfunctions). For example, a researcher interested in a dyad-year analysis of inter-state disputes can create a non-directed dyad-year data set from 1816 to 2010 in `{peacesciencer}` with the `create_dyadyears()` function, one of the aforementioned "create" function. The following would create the base data of interest to the user (i.e. all non-directed dyad-years from 1816 to 2010).
